---
title: "Class_task"
format: html
---

# Functions
Anytime you use something multiple times 
--> write a function = shorter and easy to read script!

Use Browser() to debug a function (allows to use predefined objects when running single lines of function code) 
--> write Q to exit debugging mode

## Default values
```{r}
my_age <- function(birthday, output_unit = "days") {
  # browser()
    difftime(Sys.time(), birthday, units = output_unit)
}

# if not stated otherwise, our function uses the unit "days"
my_age("1997-04-23")
my_age("1998-09-26")
```

# Task: Writing functions 
## BMI
```{r}
bmi <- function(weight, height){
  bmi <- weight/(height)^2
  bmi
}

bmi(60, 1.56)
```

## Celsius to Farenheit
```{r}
farenheit <- function(celsius){
  fh <- celsius * 9/5 + 32
  fh
}

farenheit(0)
```

## Euclidean DIstance
```{r}
ed <- function(x, y, x2, y2){
  ed <- sqrt((x2-x)^2+(y2-y)^2)
  ed
}

ed(1,1, 2,2)
```

# Prepare analysis 
In the next tasks we will look for “meet” patterns in our wild boar data. To simplify this, we will only use a subset of our wild boar data: The individuals Rosa and Sabi for the timespan 01.04.2015 - 15.04.2015. You can download the dataset here wildschwein_BE_2056.csv. Import the csv as a data.frame and filter it with the aforementioned criteria. You do not need to convert the data.frame to an sf object.

## Dataimport
```{r}
library(readr)

wildschwein <- read_delim("wildschwein_BE_2056.csv", delim = ",")

wildschwein
```

## Filter Data
```{r}
# The individuals Rosa and Sabi for the timespan 01.04.2015 - 15.04.2015

library(dplyr)

wildschwein <- wildschwein |> 
  filter(TierName %in% c("Rosa", "Sabi"))

# be explicit what time zone is entered 
from <- as.POSIXct("2015-04-01", tz= "UTC")
to <- as.POSIXct("2015-04-15", tz= "UTC")

wildschwein <- wildschwein |> 
  filter(DatetimeUTC > from, DatetimeUTC < to)
```


## Plot
```{r}
library(ggplot2)
ggplot(wildschwein, aes(E,N, colour = "TierName"))+
  geom_point() +
  coord_equal()
```

## One Dataframe with time-alligned fixes
```{r}
library(lubridate)

rounded <- wildschwein |> 
  mutate(DateTimeRound = round_date(DatetimeUTC, "15 minutes"))

sabi <- rounded |> 
  filter(TierName=="Sabi")
rosa <- rounded |> 
  filter(TierName=="Rosa")

joined <- sabi |> 
  full_join(rosa, by = "DateTimeRound", suffix = c("_sabi", "_rosa"))

# Function
ed <- function(x, y, x2, y2){
  ed <- sqrt((x2-x)^2+(y2-y)^2)
  ed
}

joined <- joined |> 
  mutate(distance = ed(E_sabi,N_sabi, E_rosa, N_rosa),
         meet = distance < 100)

plot <- ggplot(joined) + 
  geom_point(aes(E_sabi, N_sabi), color = "blue") +
  geom_point(aes(E_rosa, N_rosa), color = "red")+
  geom_segment(data = filter(joined, meet), # filter by default only takes trues
              aes(x = E_rosa, y = N_rosa, xend = E_sabi, yend= N_sabi), 
              color= "black") +
  coord_equal()

# to zoom in
library(plotly)
ggplotly(plot)

meanmeetpoints <- joined |> 
  filter(meet) |> 
  mutate(E.mean = (E_rosa + E_sabi)/2,
         N.mean= (N_rosa + E_sabi)/2)
```

## Task 6
 *Visualize data as timecube with plotly*
 
 Finally, you can nicely visualize the meeting patterns and trajectories in a Space-Time-Cube (Hägerstraand 1970) with the package plotly. There are some nice ressources available online.
```{r}
joined |> 
  filter(DateTimeRound < "2015-04-04") |> 
  plot_ly(x = ~E_rosa, y = ~N_rosa, z = ~DateTimeRound, 
          type = "scatter3d", mode = "lines") |> 
  add_trace(joined,x = ~E_sabi, y = ~N_sabi, z=~DateTimeRound) |> 
  add_markers(data= meanmeetpoints, x= ~E.mean, y= ~N.mean, z= ~DateTimeRound) |> 
  layout(scene = list(
    xaxis= list(title = "E"),
    yaxis= list(title = "N"), 
    zaxis= list(title = "Time")
         ))
```

Somehow the frame of the markers seems to differ from the one of the joined df. Nils used another dataset (wildschwein_joined) for the add_trace command, but I am not sure what the difference to joined would be... and why my plot portrays the boar traces in 2D on the E-axis... 


